Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-28 TACAS2018/paper.tex
removed some todos
nudge

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, 
starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs taken from the set $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = 
\big\{ \mat{A}\vec{x}_k +
\mat{B}\vec{u}_k \mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
\note{do we actually need to mention the safety spec here given that it is formally introduced only in Sec~\ref{sec:cof_safety}?}
%
Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs ($X_0$ and $U$, respectively), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states reached from
$X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\forall k\in [0,n-1] \, \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}), \,
\vec{u}_{k} \in U, \,   
\vec{x}_0 \in X_0
\} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and vector sums preserve convexity \note{ref}, 
the sets $X_n = \tau^n(X_0,U)$ are also convex.  
We define the \emph{$n$-reach tube} as 
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$ which we refer to as
the reach space of the system generalises the previous notions over an unbounded
time horizon (transitive closure). \note{this sentence is grammatically incorrect.}

@@

A closed-loop digital control system with observer is shown in Figure~\ref{fig:observersystem}. 
The linear feedback controller (bottom left) is characterized by a gain matrix $\mat{K}_{\langle C \rangle}$ that influences the closed-loop dynamics, 
and is driven by the observer state $\hat{\vec{x}}$.  
Moreover, $\vec{y}_e=\hat{\vec{y}} - \vec{y}$ is the output estimation error,  
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
which is amplified by a gain matrix $\mat{L}_{\langle C \rangle}$.
Note that the controller and the observer are implemented in a discrete domain 
(quantities within this domain are identified by the index $\langle C \rangle$), whereas the plant is continuous.
Figure~\ref{fig:observersystem} also contains references to $\nu_1$, $\nu_2$, $\nu_1'$, $\nu_2'$ and $\nu_t$, which
  denote noise terms that will be explained later.   
%%   With respect to noises (denoted by $\nu$), 
%% $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations.
  Lastly, $k_r$ is a reference gain required by the model.  
  Our goal in this work is to synthesise both $\mat{K}_{\langle C \rangle}$ and $\mat{L}_{\langle C \rangle}$, 
  so that specific closed-loop properties of the overall interconnection are satisfied.

@@

\note{ [add?] In this section we work with fully observed models, namely where $y = x$. }

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated\! dynamics\! for\! continuous-time\! models\! without\! inputs}\label{sec:cont_acc_no_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

The final case we analyse is that of hybrid-time models, 
where the plant dynamics evolve in continuous time while the feedback dynamics evolve in discrete time.  
A new input value is produced only at every sample time but implemented over the continuous-time plant dynamics. 
We assume that $\mat{A}$ is invertible, since otherwise the term $B_T$ below would have to be calculated by integration, which is impractical. \note{can we say something like: for simplicity and
without loss of generality, or is this assumption restrictive?}

%
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
 with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn2}, 
 and where $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$, 
 and $\mat{A}_d, \mat{B}_d$ describe the discretized plant dynamics with sample time $T_s$, 
 is the state of the model 
 \begin{align}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=\mat{K}\vec{x}_k,  \,
 kT_s \leq t < (k+1)T_s. 
 \end{align}
 \end{theorem}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}

@@

\note{Can we just mention what each $\nu_1$, $\nu_2$, $\nu_1'$, $\nu_2'$ and $\nu_t$ in Figure~\ref{fig:observersystem} denotes -- just in plain english? In sec 3.3 we say that they will be explained later but they never are.}

@@

CEGIS-OR shows promise in exploiting modularity for improved response. 
It however relies on the use of application-specific optimization, 
which requires adequate tools for exploring different types of properties.  
Our results show significant improvement in terms of scalability over standard CEGIS approaches, 
with an ability to explore complex properties, hence many synthesis tasks may be tackled with it. 
In particular we note that our synthesis loop should be targeted at 
models where the solution space is formed by connected spaces where
optimization algorithms or localized synthesis are likely to be more
efficiently than global synthesis \note{such solution spaces are unlikely classifiable a priori}. 
Finally our application to control synthesis provides a solid proof
of concept for scalable optimal control synthesis of LTI systems.

@@

\todo{preliminaries should also contain the observer model.}
\todo{why is sec 4 not considering the observer model or a system with output?}
\todo{sec 6 needs to be merged into 7.}
\todo{transient response for an observed model?}

>>>>

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, 
starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs taken from the set $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = %\big\{ 
\mat{A}\vec{x}_k + \mat{B}\vec{u}_k
 %\mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%\note{do we actually need to mention the safety spec here given that it is formally introduced only in Sec~\ref{sec:cof_safety}?}
%
%Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs ($X_0$ and $U$, respectively), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states reached from
$X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\forall k\in [0,n-1] \, \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}), \,
\vec{u}_{k} \in U, \,   
\vec{x}_0 \in X_0
\} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and vector sums preserve convexity \note{ref}, 
the sets $X_n = \tau^n(X_0,U)$ are also convex.  
We define the \emph{$n$-reach tube} as 
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, the set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$, which we refer to as
the reach space of the system, generalises the previous notions over an unbounded
time horizon (transitive closure).

@@

A closed-loop digital control system with observer is shown in Figure~\ref{fig:observersystem}. 
The linear feedback controller (bottom left) is characterized by a gain matrix $\mat{K}_{\langle C \rangle}$ that influences the closed-loop dynamics, 
and is driven by the observer state $\hat{\vec{x}}$.  
Moreover, $\vec{y}_e=\hat{\vec{y}} - \vec{y}$ is the output estimation error,  
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
which is amplified by a gain matrix $\mat{L}_{\langle C \rangle}$.
Note that the controller and the observer are implemented in a discrete domain 
(quantities within this domain are identified by the index $\langle C \rangle$), whereas the plant is continuous.
We futher model the effects of using a discrete controller as the noise signals $\nu_1$(ADC quantization), $\nu_2$ (DAC quantization), $\nu_1'$ (rounding of the observer state error),
$\nu_2'$ (rounding of the observer state), and $\nu_t$ (output delay of the digital system) 
  Lastly, $k_r$ is a reference gain required by the model.  
  Our goal in this work is to synthesise both $\mat{K}_{\langle C \rangle}$ and $\mat{L}_{\langle C \rangle}$, 
  so that specific closed-loop properties of the overall interconnection are satisfied.

@@

%\note{ [add?] In this section we work with fully observed models, namely where $y = x$. } 
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated\! dynamics\! for\! continuous-time\! models\! without\! inputs}\label{sec:cont_acc_no_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

The final case we analyse is that of hybrid-time models, 
where the plant dynamics evolve in continuous time while the feedback dynamics evolve in discrete time.  
A new input value is produced only at every sample time but implemented over the continuous-time plant dynamics. 
%
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
 with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn2}, 
 and where $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$
  \footnote{as before, when $\mat{A}$ is not invertible, we use $\left[\begin{array}{cc}\mat{A}_T&\mat{B}_T\\ 0&\mat{I}\end{array}\right]=e^{\left[\begin{array}{cc}\mat{A}&\mat{B}\\0&0\end{array}\right]}$}, 
 and $\mat{A}_d, \mat{B}_d$ describe the discretized plant dynamics with sample time $T_s$, 
 is the state of the model 
 \begin{align}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=\mat{K}\vec{x}_k,  \,
 kT_s \leq t < (k+1)T_s. 
 \end{align}
 \end{theorem}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}

@@

@@

CEGIS-OR shows promise in exploiting modularity for improved response. 
It however relies on the use of application-specific optimization, 
which requires adequate tools for exploring different types of properties.  
Our results show significant improvement in terms of scalability over standard CEGIS approaches, 
with an ability to explore complex properties, hence many synthesis tasks may be tackled with it. 
In particular we note that our synthesis loop is more efficient for
models where the solution space is formed by connected spaces where
optimization algorithms or localized synthesis are likely to be more
efficiently than global synthesis. 
Finally our application to control synthesis provides a solid proof
of concept for scalable optimal control synthesis of LTI systems.

@@
